Determinism, Complexity, and Predictability in Computer Performance

### 2014-05-30 response.tex

\section*{Response to the First Referee}

\emph{In this manuscript the authors present a way to explore the relationship
between predictability and complexity across a broad (but not complete) array of
forecast strategies such as random walk (last value), na\"ive (average value),
regression based (ARIMA) and a nonlinear method based on state space
reconstruction (LMA). They use weighted permutation entropy as a criterion/index
of predictive structure in the time series and illustrate their approach on
eight real-world datasets derived from three different systems (computer
performance traces).}

\emph{Overall the manuscript is well written and clearly understandable. The
analysis seems to be carried out carefully and the conclusions drawn are quite
reasonable. While I don't get the impression that this is a breakthrough paper I
can still recommend the manuscript for publication in PRE if the issues listed
below are addressed thoroughly.}

We thank the reviewer for their kind words and appreciate the time they put in
to reviewing our work.

@@

\noindent\emph{Major issues:}

\emph{I don't get the reason for normalizing the three methods by the random
walk method (for example in Table I). As the authors write themselves this makes
the indicator vulnerable to a bias caused by the high influence of this
(arbitrarily selected) reference method. Why not just use a non-referential
estimator, which then would also allow to compare the inherent predictability of
different systems (as it is now the values obtained for each systems are
strongly influenced by the random walk estimate on that particular system and
thus can not be compared). This will also change the results of the fitting (the
authors explicitly exclude one value because of the effect caused by the
normalization). And it is not that the random walk predictor is the one standard
method that all other methods have to compete against.}

While we recognize the drawbacks inherent in using the industry-standard error
metric MASE, we chose it over other error metrics due to some of its favorable
qualities; namely that it is scale-free and comparable across different time
series. In our opinion this made it the superior candidate despite the bias it
can introduce.

@@

\emph{In many parts the article reads almost like a review. There is a lot of
unnecessary redundancy (the entropy rate is quite low). In particular, the main
point (``The aim of the paper is not ... but ...'') is repeated far too many
times. The paper while easy to read could be shortened considerably without any
loss of information.}

Due to the broad selection of fields that this work draws upon, we thought it
necessary to provide background for each of them in case the reader is not
familiar with some subset of them. While this necessitated some redundancy,
\alert{we have made an effort to trim the fat and make the review portions more
streamlined.}

@@

\noindent\emph{Minor issues:}

\emph{Page 1: Why mention the `halting problem'? I don't see a straightforward
connection to the problem at hand. The only connection is `undecidability' but
there are many other examples for that.}

Our intent in mentioning this theoretical result was to side-step the possible
response of simply ``always use the most sophisticated prediction scheme''. This
corollary implies that no matter how sophisticated and broad a scheme we
construct, it will not be the optimal one for some time series. Thus there is a
real need for determining when a better scheme exists. \alert{We have modified
the paragraph to make this point more clear.}

\emph{Page 10: `error between the predictions and the true continuations'
$\rightarrow$ `error of the predictions with respect to the true continuations'
Figure 6 on Page 11: It would be more intuitive to turn this into a `table of
figures' with the four methods as row labels on the left and the three systems
as column labels at the top. Also: Caption `for forecast of ... and all four
prediction strategies'.}

We have modified the text as the reviewer suggested, and agree that it reads
more clearly now. \alert{We have also modified the figure as suggested.}
